Verification of OpenSSL's SHA3 Keccak function using the Software Analysis Workbench

### 2021-09-10 paper/amended.tex

\section{Amended Cryptol}\label{sec:amended}
Initially it appeared that visual inspection with Cryptol’s standard syntax was subpar.
Cryptol’s functional folds and list comprehensions do not align easily with the imperative 
and array-based NIST standard. By leveraging Cryptol’s versatility, however, we produced a couple 
utility functions to visually mimic the sequential nature of the NIST specification.

@@

\begin{figure}[h]
  \centering
\begin{verbatim}
type COL = 5 
type W = 64
type STATE_ARR = [COL][COL][W]
LIST4 = [0..COL-1]
LIST63 = [0..W-1]
pi : STATE_ARR -> STATE_ARR
pi a = for LIST4 (\y ->
          for LIST4 (\x -> 
            for LIST4 (\z -> 
              a @x @((x + 3*y) % `COL) @z)))
\end{verbatim}
\caption{Ammended Cryptol Pi definition}
\label{fig:cryptolAmmendedPi}
\end{figure}

The ‘for’ utility function follows the same intuition as the implementation in most 
imperative languages. That is, it iterates over the list provided. It is similar to the 
Array.prototype.map method in JavaScript in that it takes in a function and a list of 
arguments and returns a list of results from calling the function on every element of the 
input array. It is implemented as so:

@@

\begin{figure}[h]
  \centering
\begin{verbatim}
rc : [64] -> Bit
rc t = if (t % 255) == 0 
  then True 
  else rs !0 @0 where
    rs = [128:[8]] # [
      if i <= t % 255
        then (take`{8} ([
          (r@0)^(r@8),
          r@1,
          r@2,
          r@3, 
          (r@4)^(r@8),
          (r@5)^(r@8),
          (r@6)^(r@8),
          r@7,
          r@8])
            where r = [False] # (rs @(i - 1)))
        else rs @(i - 1)
      | i <- [1..254]]
\end{verbatim}
\caption{Cryptol RC definition}
\label{fig:cryptolRC}
\end{figure}

@@

\begin{figure}[h]
  \centering
\begin{verbatim}
rc : [64] -> Bit
rc t = if (t % 255) == 0 
  then True 
  else (while {i = 1, R = 128:[8]} //STATE
    (\state -> state.i <= t % 255) //COND
    (\state -> {                   //LOOP
      i = state.i + 1, 
      R = (take`{8} ([
        (r@0)^(r@8),
        r@1,
        r@2,
        r@3,
        (r@4)^(r@8),
        (r@5)^(r@8),
        (r@6)^(r@8),
        r@7, 
        r@8]) 
          where r = [False] # state.R)})
    ).R @0
\end{verbatim}
\caption{Ammended Cryptol RC definition}
\label{fig:cryptolAmmendedRC}
\end{figure}

>>>>

\section{Amended Cryptol}
\label{sec:amended}
Initially it appeared that visual inspection with Cryptol’s standard syntax was subpar.
The language's functional folds and list comprehensions do not align easily with the imperative 
and array-based NIST standard. By leveraging Cryptol’s versatility, however, we produced a couple 
utility functions to visually mimic the sequential nature of the NIST specification.

@@

\begin{figure}[h]
  \centering
\begin{verbatim}
type COL = 5 
type W = 64
type STATE_ARR = [COL][COL][W]
LIST4 = [0..COL-1]
LIST63 = [0..W-1]
pi : STATE_ARR -> STATE_ARR
pi a = for LIST4 (\x ->
          for LIST4 (\y -> 
            for LIST4 (\z -> 
              a @x @((x + 3*y) % `COL) @z)))
\end{verbatim}
\caption{Amended Cryptol Pi definition}
\label{fig:cryptolAmendedPi}
\end{figure}

The ‘for’ utility function follows the same intuition as the implementation in most 
imperative languages in that it iterates and runs a loop body over defined elements. 
It is similar to the Array.prototype.map method in JavaScript in that it takes in a 
function and a list of arguments and returns a list of results from calling the function 
on every element of the input array. It is implemented as so:

@@

\begin{figure}[h]
  \centering
\begin{verbatim}
rc : [64] -> Bit
rc t = if (t % 255) == 0 
  then True 
  else rs !0 @0 where
    rs = [128:[8]] # [
      if i <= t % 255
        then (take`{8} ([
          (r@0) ^ (r@8),
          r@1,
          r@2,
          r@3, 
          (r@4) ^ (r@8),
          (r@5) ^ (r@8),
          (r@6) ^ (r@8),
          r@7,
          r@8])
            where r = [False] # (rs @(i - 1)))
        else rs @(i - 1)
      | i <- [1..254]]
\end{verbatim}
\caption{Cryptol RC definition}
\label{fig:cryptolRC}
\end{figure}

@@

\begin{figure}[h]
  \centering
\begin{verbatim}
rc : [64] -> Bit
rc t = if (t % 255) == 0 
  then True 
  else (while {i = 1, R = 128:[8]} //STATE
    (\state -> state.i <= t % 255) //COND
    (\state -> {                   //LOOP
      i = state.i + 1, 
      R = (take`{8} ([
        (r@0) ^ (r@8),
        r@1,
        r@2,
        r@3,
        (r@4) ^ (r@8),
        (r@5) ^ (r@8),
        (r@6) ^ (r@8),
        r@7, 
        r@8]) 
          where r = [False] # state.R)})
    ).R @0
\end{verbatim}
\caption{Amended Cryptol RC definition}
\label{fig:cryptolamendedRC}
\end{figure}
